Component response rate variation underlies the stability of complex systems

### 2020-03-05 notebook/ms.Rmd
Remove the last figure, which will now be placed in the supporting information

**Figure 3: System correlation versus stability across different system sizes**.

@@

mtext(text = "E leading real part eigenvalue", side = 2,

@@

**Figure 5: Stability of large complex systems given $\boldsymbol{\gamma = 1}$ versus targeted $\boldsymbol{Var(\gamma)}$.** The $\ln$ number of systems that are stable across different system sizes ($S$, max $S=40$) for $C = 1$, and the proportion of systems wherein a targeted search of $\gamma$ values successfully resulted in system stability. For each $S$, 100000 complex systems are randomly generated. Stability of each complex system is tested given variation in $\gamma$ using a genetic algorithm to maximise the effect of $\gamma$ values on increasing stability, as compared to stability in an otherwise identical system in which $\gamma$ is the same for all components. Blue bars show the number of stable systems in the absence of component response rate variation, while red bars show the number of stable systems that can be generated if component response rate is varied to maximise system stability. The black line shows the proportion of systems that are stable when component response rate is targeted to increase stability, but would not be stable if $Var(\gamma) = 0$.

```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
dat <- read.csv(file = "sim_results/evolved/evo_results.csv");
dat <- dat[,-1];
plot_stable_4(dat);
```

>>>>

**Figure 3: System correlation versus stability across different system sizes**. In each panel, 10000 random complex systems $\mathbf{M} = \gamma \mathbf{A}$ are simulated for each correlation $\rho = \{-0.90, -0.85, ..., 0.85, 0.90 \}$ between off-diagonal elements $A_{ij}$ and $A_{ji}$. Lines show the expected real part of the leading eigenvalues of $\mathbf{M}$ (red squares) versus $\mathbf{A}$ (blue circles) across $\rho$, where negative values (below the dotted black line) indicate system stability. Differences between lines thereby show the effect of component response rate variation ($\gamma$) on system stability across system correlations and sizes ($S$). For all simulations, system connectance was $C = 1$. Off-diagonal elements of $\textbf{A}$ were randomly sampled from $A_{ij} \sim \mathcal{N}(0, 0.2)$, and diagonal elements were set to $-1$. Elements of $\gamma$ were sampled such that $\gamma_{i} \sim \mathcal{U}(0, 2)$, so $\sigma^{2}_{\gamma} = 1/3$.

@@

mtext(text = "E real part of leading eigenvalue", side = 2,

@@

### 2020-03-05 notebook/ms.tex
Remove the last figure, which will now be placed in the supporting information

\textbf{Figure 1: Eigenvalue distributions of random complex systems.}
Each panel shows the real (x-axis) and imaginary (y-axis) parts of
\(S =\) 400 eigenvalues from random \(S \times S\) matrices.
(\(\textbf{a}\)) A system represented by a matrix \(\mathbf{A}\), in
which all elements are sampled from a normal distribution with
\(\mu = 0\) and \(\sigma_{A} = 1/\sqrt{S}\). Points are uniformly
distributed within a circle centred at the origin with a radius of
\(\sigma_{A} \sqrt{S} = 1\). (\(\textbf{b}\)) The same system after
including variation in the response rates of \(S\) components,
represented by the diagonal matrix \(\gamma\), such that
\(\mathbf{M} = \gamma\mathbf{A}\). Elements of \(\gamma\) are randomly
sampled from a uniform distribution from \(\min = 0\) to \(\max = 2\).
Eigenvalues of \(\mathbf{M}\) are then distributed non-uniformly within
a circle centred at the origin with a radius of
\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.21.
(\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed
from the same parameters as in \(\textbf{a}\), except with diagonal
values of \(-1\). (\(\textbf{d}\)) The same system \(\textbf{c}\) after
including variation in component response rates, sampled from
\(\mathcal{U}(0, 2)\) as in \(\textbf{b}\).

@@

\textbf{Figure 3: System correlation versus stability across different
system sizes}.

@@

\textbf{Figure 5: Stability of large complex systems given
\(\boldsymbol{\gamma = 1}\) versus targeted
\(\boldsymbol{Var(\gamma)}\).} The \(\ln\) number of systems that are
stable across different system sizes (\(S\), max \(S=40\)) for
\(C = 1\), and the proportion of systems wherein a targeted search of
\(\gamma\) values successfully resulted in system stability. For each
\(S\), 100000 complex systems are randomly generated. Stability of each
complex system is tested given variation in \(\gamma\) using a genetic
algorithm to maximise the effect of \(\gamma\) values on increasing
stability, as compared to stability in an otherwise identical system in
which \(\gamma\) is the same for all components. Blue bars show the
number of stable systems in the absence of component response rate
variation, while red bars show the number of stable systems that can be
generated if component response rate is varied to maximise system
stability. The black line shows the proportion of systems that are
stable when component response rate is targeted to increase stability,
but would not be stable if \(Var(\gamma) = 0\).

\includegraphics{ms_files/figure-latex/unnamed-chunk-16-1.pdf}

>>>>

\textbf{Figure 1: Eigenvalue distributions of random complex systems.}
Each panel shows the real (x-axis) and imaginary (y-axis) parts of
\(S =\) 400 eigenvalues from random \(S \times S\) matrices.
(\(\textbf{a}\)) A system represented by a matrix \(\mathbf{A}\), in
which all elements are sampled from a normal distribution with
\(\mu = 0\) and \(\sigma_{A} = 1/\sqrt{S}\). Points are uniformly
distributed within a circle centred at the origin with a radius of
\(\sigma_{A} \sqrt{S} = 1\). (\(\textbf{b}\)) The same system after
including variation in the response rates of \(S\) components,
represented by the diagonal matrix \(\gamma\), such that
\(\mathbf{M} = \gamma\mathbf{A}\). Elements of \(\gamma\) are randomly
sampled from a uniform distribution from \(\min = 0\) to \(\max = 2\).
Eigenvalues of \(\mathbf{M}\) are then distributed non-uniformly within
a circle centred at the origin with a radius of
\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.14.
(\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed
from the same parameters as in \(\textbf{a}\), except with diagonal
values of \(-1\). (\(\textbf{d}\)) The same system \(\textbf{c}\) after
including variation in component response rates, sampled from
\(\mathcal{U}(0, 2)\) as in \(\textbf{b}\).

@@

\textbf{Figure 3: System correlation versus stability across different
system sizes}. In each panel, 10000 random complex systems
\(\mathbf{M} = \gamma \mathbf{A}\) are simulated for each correlation
\(\rho = \{-0.90, -0.85, ..., 0.85, 0.90 \}\) between off-diagonal
elements \(A_{ij}\) and \(A_{ji}\). Lines show the expected real part of
the leading eigenvalues of \(\mathbf{M}\) (red squares) versus
\(\mathbf{A}\) (blue circles) across \(\rho\), where negative values
(below the dotted black line) indicate system stability. Differences
between lines thereby show the effect of component response rate
variation (\(\gamma\)) on system stability across system correlations
and sizes (\(S\)). For all simulations, system connectance was
\(C = 1\). Off-diagonal elements of \(\textbf{A}\) were randomly sampled
from \(A_{ij} \sim \mathcal{N}(0, 0.2)\), and diagonal elements were set
to \(-1\). Elements of \(\gamma\) were sampled such that
\(\gamma_{i} \sim \mathcal{U}(0, 2)\), so \(\sigma^{2}_{\gamma} = 1/3\).

@@
